% Draft #1

\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[super]{nth}
\usepackage[all]{nowidow}
\begin{document}

	\centerline{\large Computer Science Project Proposal}
	\vspace{0.4in}
	\centerline{\Large Building a Gradual Type System for JavaScript }
	\vspace{0.4in}
	\centerline{\large Christopher J.O. Little, Emmanuel College}
	\vspace{0.3in}
	\centerline{\large \today}
	\vspace{0.3in}

	\noindent{\bf Project Supervisor:} Dr Kathryn E. Gray
	\vspace{0.2in}

	\noindent{\bf Director of Studies:} Dr Jonathan Hayman
	\vspace{0.2in}\noindent 
	
	\noindent{\bf Project Overseers:} Dr~Markus~Kuhn \& Dr~Neal~Lathia


	\section{Introduction}\label{prop-introduction}

		Static type systems can offer a number of advantages over their dynamic
		counterparts, including earlier detection of certain programming
		errors, better development tools and potential for compiler
		optimisations. Proponents of dynamically typed languages may argue that
		static type systems are too inflexible and interfere with the natural
		thought processes of the developer.
		
		JavaScript is a dynamic language with little to no built-in support for
		types. A few projects have attempted to integrate optional explicit
		type annotations to dialects of ECMAScript (JScript.NET, ActionScript
		3.0, ECMAScript 4 and more recently TypeScript), but the requirement
		for explicit type annotations limits the utility of these for existing
		JavaScript projects. Often this requirement is relaxed such that type
		annotations are optional, but in these cases it is not evident to the
		developer which parts of the codebase are type-safe, and which have not
		been checked.

		Both of these problems could be resolved with a gradual type inferring
		system. Gradual typing is a modern technique allowing safe
		interoperability between typed and untyped portions of a single
		program. The technique allows us to guarantee that only dynamic
		properties of the code are able to cause runtime errors, retaining the
		utility of statically type-checking part of the code.
		\cite{felleisen06,findler07}

		Using type inference means no extensions to the language are required,
		such that the system could be used to retrospectively analyse existing
		code. Due to the highly dynamic nature of JavaScript, it will not
		always be possible to statically determine the type of an expression.
		This is where the gradual typing nature of the system becomes critical
		- we can safely assert that not only are the statically analysed
		portions of the program free of type errors, they are also guaranteed
		to remain so even after interaction with potentially untyped code.

	\section{Starting Point}\label{prop-starting-point}

		I am comfortable writing programs in JavaScript and familiar with many
		of the semantic quirks of the language. Many of the concepts required
		for the execution of this project were introduced in the Part IB
		courses Compiler Construction and Semantics. The Part II courses on
		Optimising Compilers and Types will also be useful.

		I have already done some preparatory reading in this area, on types in
		general \cite{types1}, on the design and semantics of JavaScript
		\cite{js1,js2}, and specifically on gradual typing
		\cite{gradual1,gradual2} and on TypeScript \cite{typescript}, a recent
		project by Microsoft extending the JavaScript language with type
		annotations.

		For the implementation of the project, it will be necessary to
		construct an abstract syntax tree from JavaScript source, and, after
		modification of the tree, reconstruct the JavaScript source code. Since
		the parsing and reconstruction of JavaScript is not of particular
		relevance to the core focus of this project, and in order to integrate
		with existing solutions in the world of JavaScript, I will use an
		existing system for this - UglifyJS.

		UglifyJS is a tool for minification and compression of JavaScript code,
		itself written in JavaScript. It uses a recursive descent parser to
		construct an abstract syntax tree from JavaScript source code, then
		modifies the tree (renaming identifiers and similar) before
		reconstructing minimal source code.
		 % is this really resursive descent?!

	\section{Resources}\label{prop-resources}

		For this project I will mainly use my own machines - either a desktop
		(Ubuntu 14.04, 3.6GHz i7, 8GB RAM, 2x 1TB HDD) or laptop (Windows 8.1,
		1.8GHz i7, 8GB RAM, 1TB HDD) as convenient. I will use Git for version
		control, with remotes set up on Github and on a VPS hosted by BHost for
		backup purposes. I require no additional special resources.
				
	\section{Work to be Done}\label{prop-work-to-be-done}

		The project can be broken into the following components:

		\begin{enumerate}\def\labelenumi{\arabic{enumi}.}
		 \item Definition of a formal type system along with an operational
		   semantics for the subset of JavaScript which is supported. The
		   system will have the following properties to prove:   
			
			\emph{Progress}: If an expression is well-typed, then either it can
			reduce further to a new expression, or it is a value (and cannot
			reduce further). 

			\emph{Type Preservation}: If an expression $e$ is well-typed and
			reduces in some number of steps to another expression, then that
			expression is also well-typed.
			
			\emph{Decidability of Type Checking}: It is decidable for any
			expression in the language whether or not it has a given type.
			
			\emph{Type inference}: It is possible to find a type for any
			expression in the language, or to show that no such type exists.

		 In addition, it will be necessary to prove that any runtime errors
		 which do arise have been caused by untyped code.
		 
		 \item Implementation of a static type checker for the above type
		   system. This will need to include a type inference algorithm in
		   order to make any deductions about the types of JavaScript objects.
		   This algorithm could be based upon a variation of the Hindley-Milner
		   algorithm.
		 
		 \item Implementation of a source-to-source compiler to implement
		   gradual typing, allowing code which has had its type checked by the
		   above type checker to safely interface with unchecked code. The
		   compiler will inject two kinds of wrappers around objects to achieve
		   this. Guard wrappers protect a well-typed object from untyped code
		   by performing dynamic type checks on parameters passed to its
		   methods, or creating a mimic wrapper for the parameter. It will also
		   create guard wrappers for return values (which are well-typed).
		   Mimic wrappers protect well-typed code from an untyped object by
		   checking for the existence of the object's properties and methods
		   when these are requested, and checking return types or creating a
		   mimic wrapper around the return value when this is not possible.
		   They will also create guard wrappers around parameters (which are
		   well-typed and entering untyped code). The operation of this
		   compiler will be validated by running it on a series of test cases,
		   and also on some existing production code. These test cases will be
		   written alongside the type checker.

		\end{enumerate}

	\section{Success Criteria}\label{prop-success-criteria}

		The project will be considered a success if the following points have been achieved:

		\begin{enumerate} 
			\item The design of an operational semantics and type system for a
			  subset of JavaScript including: 

			\begin{enumerate}

				\item Primitives, constants and object literals  
				\item The following control flow statements   
				% potentially also others wouldn't be too hard from https://developer.mozilla.org/en-US/docs/Web/JavaScript/Guide/Statements 
				% these were just the "main" ones   
				\begin{enumerate}    
					\item seq ( ... ; ... )   
					\item block ( \{ ... \} )    
					\item if ... else    
					\item while    
					\item for     
					\item continue    
					\item break    
				\end{enumerate}  
				\item Assignment via =, and all shorthand assignment operators (such as +=)  
				\item Arithmetic operators +, -, *, /  
				\item Logical operators !, \&\&, ||  
				\item String concatenation with +  
				\item Function definition 
			\end{enumerate} 
			
			\item An implementation of a type checker for a subset of
			  JavaScript defined by the list above. 

			\item An implementation of a source-to-source compiler injecting
			  code to provide type-safe wrappers around dynamic type-unsafe
			  objects.
		  \end{enumerate}

		Evaluation of the design of my type system will be achieved through a
		formal proof of soundness. Evaluation of the implementation will be
		through testing to monitor performance and memory overhead.

	\section{Possible Extensions}\label{prop-possible-extensions}

		Arrays in JavaScript pose a difficult problem, since the types of their
		contents are unrestricted. My project will thus be limited to programs
		which do not include arrays, but it would in principle be possible to
		extend the type system to handle them. This extension would also allow
		handling of variadic functions, which access their actual parameters
		through the \texttt{arguments} array.

		A common idiom in JavaScript is to initialise an object with certain
		properties and methods, but then to modify the object later on, either
		adding or removing properties. This kind of behaviour is difficult to
		support in a static type system, but could be supported with gradual
		typing.

		Objects in JavaScript can be created by a literal object initialiser,
		or in a more object-oriented style with the \texttt{new} keyword. Using
		\texttt{new} with a constructor creates a new object following an
		inheritance chain and binds the special \texttt{this} variable to the
		new object. The inheritance involved in this method complicates typing,
		but this would be a valuable extension to my project, as \texttt{new}
		is a common way to initialise objects.

		It would also be informative to run the project on larger JavaScript
		programs, in order to measure the impact of the compiler on performance
		and memory in the presence of a larger codebase. An extension might
		thus be to run the compiler on at least a portion of Mozilla's
		SunSpider JavaScript benchmark series.

	\section{Timetable}\label{prop-timetable}

		\subsection*{Weeks 1 \& 2 (\nth{24} Oct -- \nth{6} Nov)}
		Continue learning about techniques used in gradual typing, and
		implementations of Hindley-Milner. Prepare an initial design for the
		type system such that most of the core features of my subset of
		JavaScript have a defined operational semantics and type judgement.

		\subsection*{Weeks 3 \& 4 (\nth{7} Nov -- \nth{20} Nov)}
		Create a trivial implementation of the type checker which simply
		rejects everything, and a trivial compiler which generates an AST and
		re-generates source code from it. Complete type system if required.

		\subsection*{Weeks 5 \& 6 (\nth{21} Nov -- \nth{4} Dec)}
		Extend the type checker and compiler allowing primitives and constants.
		Prove safety and write test cases for these new cases. Begin extension
		for variable initialisation and assignment.

		\subsection*{Weeks 7 \& 8 (\nth{5} Dec -- \nth{18} Dec)}
		Complete extension for assignment, and add extension for object
		literals. Prove safety and write test cases for this extension.

		\subsection*{Weeks 9 \& 10 (\nth{19} Dec -- \nth{1} Jan)}
		Extend to support all control flow statements. Prove safety and write
		test cases for this  extension.

		\subsection*{Weeks 11 \& 12 (\nth{2} Jan -- \nth{15} Jan)}
		Ensure development is on schedule, and prepare progress report. 

		\subsection*{Weeks 13 \& 14 (\nth{16} Jan -- \nth{29} Jan)}
		Extend the type checker and compiler for function definitions and
		arithmetic and logical operators. Prove safety and write test cases for
		this extension. By this point, all core features of the project should
		be implemented.

		\subsection*{Weeks 15 \& 16 (\nth{30} Jan -- \nth{12} Feb)}
		Finalise development and fix any bugs. Complete test harness.

		\subsection*{Weeks 17 \& 18 (\nth{13} Feb -- \nth{26} Feb)}
		Evaluate performance and additional memory overhead of compiled
		JavaScript. Complete the proofs of soundness if not yet done.

		\subsection*{Weeks 19 -- 23 (\nth{27} Feb -- \nth{26} Mar)}
		Write dissertation and refine after feedback before submission. Prepare
		plan for tackling any extension work.
		
		\subsection*{Weeks 24 -- 28 (\nth{27} Mar -- \nth{23} Apr)}
		Extend project as far as possible, and update dissertation as
		appropriate.

	\begin{thebibliography}{9}

		\bibitem{felleisen06}
		S. Tobin-Hochstadt and M. Felleisen:
		\emph{Interlanguage Migration: From Scripts to Programs}.
		In: Proc. of ACM Dynamic Languages Symposium. (2006)

		\bibitem{findler07} P. Wadler and R.B. Findler: 
		\emph{Well-typed programs can't be blamed}
		In: Proc. ACM Workshop on Scheme and Functional Programming (2007)

		\bibitem{js1} A. Guha, C. Saftoiu, and S. Krishnamurthi: 
		\emph{The Essence of JavaScript}
		In: ECOOP (2010)

		\bibitem{js2} G. Richards, S. Lebresne, B. Burg and J. Vitek:
		\emph{An Analysis of the Dynamic Behavior of JavaScript Programs}
		In: Proc. ACM SIGPLAN Conference on Programming Language Design and Implementation (2010)

		\bibitem{types1} B. Pierce:
		\emph{Types and Programming Languages} (2002)

		\bibitem{gradual1} K. Gray: 
		\emph{Safe Cross-Language Inheritance}
		In: ECOOP (2008)

		\bibitem{gradual2} E. Meijer and P. Drayton
		\emph{Static Typing Where Possible, Dynamic Typing When Needed: The End of the Cold War Between Programming Languages}
		In: OOPSLA Workshop on Revival of Dynamic Languages (2004)

		\bibitem{typescript} G. Bierman, M. Abadi and M. Torgersen: 
		\emph{Understanding TypeScript}
		In: ECOOP (2014)

	\end{thebibliography}


\end{document}
